---
title: "**Task 2 - Global Terrorism Analysis**"
output: html_document
header-includes:
    - \usepackage{sectsty}
    - \allsectionsfont{\color{cyan}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


<style>
#header{
color: #720e07;
font-size: 30px;
background-color: #eecf6d;
font-family: "Times New Roman";
text-align: center;
}
h3{
text-align: left;
font-family: "Times New Roman";
color:#720e07;
}
body{
font-family: "Times New Roman";
font-size: 20px;
background-color: #ffffff;
}
h1,h2,h4{
text-align: center;
color: #720e07;
}
.center {
  display: block;
  margin-left: auto;
  margin-right: auto;
  width: 50%;
}
.intro{
color: #720e07;
text-align: center;
}
.list{
text-align: left
}
</style>




<h3 class="intro"> **The Sparks Foundation** </h3>
#### **Data Science & Business Analytics Internship - February 2022**
#### **Ahmed Yasser Disoky Dar**

<img src = logo.png style="width:300px;" class = "center">




### **About the Dataset**

The dataset **_Global Terrorism_** contains data about crimes that took place from **1970** to **2017** in **205** countries from all over the globe. The dataset consists of **181,691** rows and **135** columns. The objective of this report is to perform an Explanatory Data Analysis on this data, answer some business questions and find interesting facts. 


<br>

### **Importing The Necessary Libraries** 
We import the relevant libraries for the analysis such as `library(gapminder)` & `library(dplyr)`.
```{r include=FALSE}
library(gapminder)
library(dplyr)
```

<br>

### **Loading the Dataset & Exploring the Structure and Content**
```{r Loading Data, echo=TRUE}
df <- read.csv("terrorism.csv", stringsAsFactors = TRUE)
glimpse(df)
```

<br>

Selecting only the columns we are interested in from the dataframe.

```{r Selecting Data}
df<- df %>% 
  select(ï..eventid:city, success:targsubtype1_txt, gname, motive:weapsubtype1_txt, nkill:property, propextent_txt, propvalue)
glimpse(df)
```

<br>

Renaming Dataset Columns for More Clear & Descriptive Labels.

```{r Renaming Data}
df <- df %>%
  rename(event_id = ï..eventid, year=iyear, month=imonth, day=iday, country=country_txt, region=region_txt, state=provstate, success_crimes=success, suicide_crimes=suicide, attack_type=attacktype1_txt, target_type=targtype1_txt, target_subtype=targsubtype1_txt, gang_name=gname, weapon_type=weaptype1_txt, weapon_subtype=weapsubtype1_txt, killed=nkill, wounded=nwound, estimated_property_range=propextent_txt, damaged_property_value=propvalue)

glimpse(df)
```

<br>

The first 5 rows of the dataset
```{r First Five}
head(df, 5)
```

<br>
Looking at the structure of the data
```{r Structure}
str(df)
```

<br>

### **Data Cleaning**
During this process, we check for missing, duplicated, or irrelevant data in our dataset to increase the potential of high-quality analysis.

<br>

First, we check for any missing values
```{r Check Null}
colSums(is.na(df))
```

Replacing Null values with Zeros
```{r Replace Null}
df <- df%>%
  mutate(killed = ifelse(is.na(killed), 0, killed)) %>%
  mutate(wounded = ifelse(is.na(wounded), 0, wounded)) %>%
  mutate(property = ifelse(is.na(property), 0, property))

sum(is.na(df$killed))
sum(is.na(df$wounded))
sum(is.na(df$property))
```

<br>

Checking whether there are any duplicated rows.
```{r Check Duplicates} 
sum(duplicated(df))
``` 

Dropping duplicates
```{r Dropp Duplicates}
df <- distinct(df)
```

<br> 
Checking the values of both `month` and `day` columns.
```{r Check Columns}
unique(df$month)
unique(df$day)
```

Eliminating rows that have 0 either in the month or day columns as they won't be relevant for the analyses.
```{r Removث Zero}
df <- df %>%
  filter(month > 0, day > 0)
```

<br>

Checking the Structure of the Cleaned Dataset.
```{r Check Structure}
str(df)
```

<br>

We export the cleaned dataset as a CSV file to be further used in the visualization phase.
```{r}
write.csv(df, "Cleaned_df.csv", row.names = TRUE)
```
<br>




### **Explanatory Data Analysis**

_a. Total number of crimes for each country._
```{r crimes_country}
df %>%
  group_by(country) %>%
  count(country, sort=TRUE)
```

b. Total number of crimes for each region.
```{r crimes_region}
df %>%
  group_by(region) %>%
  count(region, sort=TRUE)
```

c. Total crimes by each gang. 
```{r}
df %>%
  group_by(gang_name) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count)) 
```

d. Total kills and wounds by each gang.
```{r}
df %>%
  group_by(gang_name) %>%
  summarize(Kills = sum(killed), Wounds = sum(wounded)) %>%
  arrange(desc(Kills, Wounds))
```

e. Kills and wounds by each country.
```{r}
df %>%
  group_by(country) %>%
  summarize(Kills = sum(killed), Wounds = sum(wounded)) %>%
  arrange(desc(Kills), desc(Wounds))
```

f. Hottest Years.
```{r}
df %>%
  group_by(year) %>%
  summarize(No_of_Crimes = n())%>%
  arrange(desc(No_of_Crimes))

```

g. Number of crimes committed by a weapon type.
```{r}
df %>%
  group_by(weapon_type) %>%
  summarize(Crimes = n()) %>%
  arrange(desc(Crimes))
```

h. Top targets.
```{r}
df %>%
  group_by(target_type) %>%
  summarize(Target = n()) %>%
  arrange(desc(Target))
```
i. Top Sub-targets by attackers
```{r}
df %>%
  group_by(target_subtype) %>%
  summarize(Target_Subtype = n()) %>%
  arrange(desc(Target_Subtype))
```

j. Most attack types committed.
```{r}
df %>%
  group_by(attack_type) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))
```

k. Safest countries 
```{r}
df %>%
  group_by(country) %>%
  summarize(Crimes = n())%>%
  arrange(Crimes)
```

l. Suicide crimes occurred in each country.
```{r}
df %>%
  group_by(country) %>%
  summarise(Crimes = sum(suicide_crimes)) %>%
  arrange(desc(Crimes))

```

m. Estimated property damages for each country.
```{r}
df %>%
  group_by(country) %>%
  summarize(Total = mean(damaged_property_value)) %>%
  mutate(Total = as.integer(Total)) %>%
  arrange(desc(Total))
```

n. Classifying the range amount of damages.
```{r}
df %>%
  group_by(estimated_property_range) %>%
  filter(estimated_property_range != "") %>%
  summarise(Total = n())%>%
  arrange(desc(Total))

```


<br>

### **Data Visualization is performed using Power BI**

<br>
<img src = powerbi_logo.png style="width:300px;" class="center">